Stability and Performance of Coalitions of Prosumers Through Diversification in the Smart Grid

### 2017-05-10 Appendix.tex
latest changes

%\onecolumn
\appendix

%\twocolumn
\subsection{Algorithms}

@@

\begin{figure}[b]
\includegraphics[scale=.5]{./figs/figure_7}
\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
\label{fig:search}
\end{figure}

@@

@@

Figure \ref{fig:notations} presents all the notations used in the article.

@@

@@

\subsubsection{Wind power curve}
Power curves are functions that, for a given type of generator, map some input quantity to the output power produced. For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007, Piedallu2008}. For the wind turbines, the power curve can be specified by 4 values :

@@

\subsubsection{Solar power curve}
The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data. As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007, Piedallu2008}. This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps. As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007, Piedallu2008} for more details) :

\begin{equation}
\left\{ \begin{array}{lll}
			\Psi{real}(t) = \Psi_{perfect}(t) \eta(t) \\
			\eta(t) = 1-0.75 \left( \dfrac{N(t)}{8} \right)^{3.4}
\end{array} \right.
\end{equation}
where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t. Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :

\begin{equation}
\mathcal{F}_{PV}(\Psi{real}(t)) =  S_{PV} \Psi{real}(t)  e_{PV}
\end{equation}
where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency. The very simple form of this power curve is due to some simplifications in order not to overload the model. For instance, it does not take into account angles and orientations degradations. These could be incorporated if needed by changing the power curve in the simulations.

@@

\begin{equation}
P_{i}^{D}(t) = \mathcal{F}_{i}^{heat}(\tau(t), t) + \mathcal{F}_{i}^{elec}(t)
\end{equation}
where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day. In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification. By using thermodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :

\begin{equation}
\mathcal{F}_{i}^{heat}(\tau(t), t) = \dfrac{B_{i}}{R_{i}} \left[ T_{i} - \tau(t) \right]
\end{equation}
where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance. We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers. We also denote by $ \omega_{i}(t) = \{ \omega_{i}(t_{0}),...,\omega_{i}(t_{24}) \} $ the vector of the average fraction of $ \Omega_{i} $ used for each hour. We can therefore write :

\begin{equation}
\mathcal{F}_{i}^{elec}(t) = \Omega_{i} ( \omega_{i}(t) + \epsilon )
\end{equation}
where $ \epsilon $ is a noise term. The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors. Business or residential areas for instance can be easily distinguished with this kind of model.

\subsection{Resilience algorithm}
\label{appendix_resilience}
The following algorithm (see algorithm \ref{alg:algo4}) explains the process through which the resilience of a coalition structure is explored. For a given coalition structure, we consider random failures of physical components of the agents. This results in the agent being unable to support its coalition. Therefore, the agent is simply removed from the coalition, which impacts the coalition's production. Eventually, after some number of agent failures, some coalitions might not be sufficiently stable and productive to fulfill the market's constraints. In this case, the coalition fails, and is removed from the market.
\begin{algorithm}
	\KwData{$ CS = \left\{ S_1,S_2,...,S_k \right\} $ Coalition structure \;}

@@

@@

Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function. For some range of values, the utility might not even be a concave function of the coalition size. Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, the algorithm \ref{alg:algo1} will start with $ N_{COAL} $ seeds and increase their sizes as long as the utilities are improved. The order through which coalitions incorporate new prosumers is randomized but leads generally to relatively balanced coalitions. We denote by $ \bar{N} $ the desired approximate final size of the coalitions, i.e we target a utility function that is concave in the coalition size with a maximum at $ \bar{N} $ (in the article, $ \bar{N} $ is fixed at $\frac{N}{N_{COAL}}$). We propose to obtain such utility function through tunning $\alpha $. Ideally, we seek to express $ \alpha $ as a function of $ \bar{N} $ and the agents power distributions and correlations.
\\
In the general case, neither the power distributions nor the correlation structure have special forms which makes obtaining an analytical expression for $ \alpha $ difficult. Therefore we use two simplifying assumptions in order to have an estimate for $ \alpha $. First, we approximate the power distributions by normal distributions, such that the optimal contract value can be written in the following way :

@@

With :

@@

Therefore, we can write the utility function as :

@@

And the utility function becomes :

@@

$ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $. Depending on $ \alpha $, the utility function can have very different shapes.
\\
We select $\alpha^{\star} $ such that :

@@

Solving these equations yields the following expression :

@@

\bibliographystyle{IEEEtran}  
\bibliography{Article}
\end{document}

>>>>

\appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%            APPENDIX A : ALGORITHMS
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Algorithms}

@@

\begin{figure}[b]
\includegraphics[scale=.5]{figure_8.eps}
\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
\label{fig:search}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%            APPENDIX B : NOTATIONS
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\noindent Figure \ref{fig:notations} presents all the notations used in the article.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%            APPENDIX C : NET PRODUCTION SERIES
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsubsection{Wind power curve}
Power curves are functions that, for a given type of generator, map some input quantity to the output power produced. For wind-turbines and solar arrays these functions are well studied and approximations have been proposed \cite{Lydia2014} \cite{Piedallu2007}, \cite{Piedallu2008}. For the wind turbines, the power curve can be specified by 4 values :

@@

\subsubsection{Solar power curve}
The input quantity desired for our power curve model for solar arrays is a radiance in $ W.m^{-2} $, which can be difficult to find in weather station available data. As we mainly collected nebulosity series, we used the Helios model described in \cite{Piedallu2007}, \cite{Piedallu2008}. This model enabled us to compute perfect (clear blue sky situation) solar radiances at some specific locations on earth and at given timestamps. As nebulosity is a measure of the sky cloudiness, we can use the nebulosity series as degradation factors on the clear blue sky model (see \cite{Piedallu2007}, \cite{Piedallu2008} for more details) :

\begin{equation}
\left\{ \begin{array}{lll}
			\Psi{real}(t) = \Psi_{perfect}(t) \eta(t) \\
			\eta(t) = 1-0.75 \left( \dfrac{N(t)}{8} \right)^{3.4}
\end{array} \right.
\end{equation}

\noindent where $ \Psi_{perfect}(t) $ and $ \Psi{real}(t) $ are respectively the clear blue sky and real radiances at time t, $ \eta(t) $ is the degradation factor at time t, and $ N(t) $ is the nebulosity index at time t. Once we have input data in the forms of radiances, we compute the production of a solar array with the following simplified power curve :

\begin{equation}
\mathcal{F}_{PV}(\Psi{real}(t)) =  S_{PV} \Psi{real}(t)  e_{PV}
\end{equation}

\noindent where $ S_{PV} $ is the surface of the array, and $ e_{PV} $ is its efficiency. The very simple form of this power curve is due to some simplifications in order not to overload the model. For instance, it does not take into account angles and orientations degradations. These could be incorporated if needed by changing the power curve in the simulations.

@@

\begin{equation}
P_{i}^{D}(t) = \mathcal{F}_{i}^{heat}(\tau(t), t) + \mathcal{F}_{i}^{elec}(t)
\end{equation}

\noindent where $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ is the power curve that maps the temperature to a heating consumption, and $ \mathcal{F}_{i}^{elec}(t) $ computes the consumption of agent i (other than heating) at a given hour of the day. In the simulation, all agents have a desired inside temperature $ T_{i} $, supposed to be a constant for simplification. By using thermodynamic laws $ \mathcal{F}_{i}^{heat}(\tau(t), t) $ can be approximated by :

\begin{equation}
\mathcal{F}_{i}^{heat}(\tau(t), t) = \dfrac{B_{i}}{R_{i}} \left[ T_{i} - \tau(t) \right]
\end{equation}

\noindent where $ B_{i} $ is the surface of thermal exchanges for agent i and $ R_{i} $ is their thermal resistance. We denote by $ \Omega_{i} $ the maximum consumption possible for agent i, which is basically the sum of all its appliances powers. We also denote by $ \omega_{i}(t) = \{ \omega_{i}(t_{0}),...,\omega_{i}(t_{24}) \} $ the vector of the average fraction of $ \Omega_{i} $ used for each hour. We can therefore write :

\begin{equation}
\mathcal{F}_{i}^{elec}(t) = \Omega_{i} ( \omega_{i}(t) + \epsilon )
\end{equation}

\noindent where $ \epsilon $ is a noise term. The vector $ \omega_{i}(t) $ enables us to easily differentiate agent consumption behaviors. Business or residential areas for instance can be easily distinguished with this kind of model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%            APPENDIX D : RESILIENCE ALGORITHM
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Resilience algorithm}
\label{appendix_resilience}

The following algorithm (see algorithm \ref{alg:algo4}) explains the process through which the resilience of a coalition structure is explored. For a given coalition structure, we consider random failures of physical components of the agents. This results in the agent being unable to support its coalition. Therefore, the agent is simply removed from the coalition, which impacts the coalition's production. Eventually, after some number of agent failures, some coalitions might not be sufficiently stable and productive to fulfill the market's constraints. In this case, the coalition fails, and is removed from the market.

\begin{algorithm}
	\KwData{$ CS = \left\{ S_1,S_2,...,S_k \right\} $ Coalition structure \;}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%%            APPENDIX E : UTILITY FUNCTION
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

Knowing what value to use for $ \alpha $ is not trivial and is important since it influences the shape of the utility function. For some range of values, the utility might not even be a concave function of the coalition size. Given a situation with N agents and a desired number of coalitions $ N_{COAL}$, the algorithm \ref{alg:algo1} will start with $ N_{COAL} $ seeds and increase their sizes as long as the utilities are improved. The order through which coalitions incorporate new prosumers is randomized but leads generally to relatively balanced coalitions. We denote by $ \bar{N} $ the desired approximate final size of the coalitions, i.e we target a utility function that is concave in the coalition size with a maximum at $ \bar{N} $ (in the article, $ \bar{N} $ is fixed at $\frac{N}{N_{COAL}}$). We propose to obtain such utility function through tunning $\alpha $. Ideally, we seek to express $ \alpha $ as a function of $ \bar{N} $ and the agents power distributions and correlations.

In the general case, neither the power distributions nor the correlation structure have special forms which makes obtaining an analytical expression for $ \alpha $ difficult. Therefore we use two simplifying assumptions in order to have an estimate for $ \alpha $. First, we approximate the power distributions by normal distributions, such that the optimal contract value can be written in the following way :

@@

\noindent With :

@@

\noindent Therefore, we can write the utility function as :

@@

\noindent And the utility function becomes :

@@

\noindent $ \bar{ \mu },\ \bar{ \sigma },\ \bar{ \rho }, and\ P^{MAX} $ are all constants in these very particular conditions, such that U is a function of the coalition size with a parameter $ \alpha $. Depending on $ \alpha $, the utility function can have very different shapes.

\noindent We select $\alpha^{\star} $ such that :

@@

\noindent Solving these equations yields the following expression :

@@

\bibliographystyle{IEEEtran}  
\bibliography{Appendix}
\end{document}
